---
title: 'DATA 605: Assignment 2'
author: "MUSA T. GANIYU"
date: "September 5, 2016"
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
  tidy: yes
  word_document: default
---
> Question 1.1:

+ Show that $A^TA\quad \neq \quad AA^T$ in general.


* Solution:

Let $A\quad =\quad \begin{matrix} a & b & c \\ d & e & f \\ g & h & i \end{matrix}$

and $A^T\quad =\quad \begin{matrix} a & d & g \\ b & e & h \\ c & f & i \end{matrix}$


$AA^T\quad =\quad \begin{matrix} a*a+b*b+c*c & d*a+e*b+f*c & g*a+h*b+i*c \\ a*d+b*e+c*f & d*d+e*e+f*f & g*d+h*e+i*f \\ a*g+b*h+c+i & d*g+h*e+i*f & g*g+h*h+i*i \end{matrix}$


where $A^TA$ would not be equal to $AA^T$. Example below shows that $A^TA\quad \neq \quad AA^T$

```{r}



A <- matrix(c(1:9), nrow=3, byrow=T)
A
t(A)

ATA <- A%*%t(A)

AAT <- t(A)%*%A

ATA
AAT

ATA==AAT

```

> Question 1.2

+ For a special type of square matrix $A^TA\quad = \quad AA^T$. Under what conditions could this be true?

* Solution: 

Consider the 2x2 identity matrix and matrix A2 below, we can deduce that the non-negative/zero of the resulting output make the matrix of  $A^TA\quad = \quad AA^T$ symmetric.
i.e if $A^TA\quad = I$ and $\quad AA^T = I$, then they are consider to be equal or symmetric.

```{R}
I <- matrix(c(1,0,0,1), nrow = 2, byrow = T)
I

A2 <- matrix(c(3,2,2,3), nrow=2, byrow=T)
A2

ATA2 <- t(A2) %*% A2
ATA2

AAT2 <- A2 %*% t(A2)
AAT2

ATA2==AAT2

```

> Question 2: Write  an  R  function  to  factorize  a  square  matrix  A  into  LU  or  LDU,  whichever  you
prefer

+ Solution:

Let the above matrix be a square matrix i.e 3x3 (mxm) dimension, then its function is as follows:
 

```{r}
if (nrow(A) == 3) {

  E21 <- -A[2,1] / A[1,1]
  matrix_E21 <- matrix(c(1,0,0,E21,1,0,0,0,1), nrow=3, byrow = T)
  A2 <- matrix_E21 %*% A
  
  E31 <- -A2[3,1] / A2[1,1]
  matrix_E31 <- matrix(c(1,0,0,0,1,0,E31,0,1), nrow=3, byrow = T)
  A3 <- matrix_E31 %*% A2
  
  E32 <- -A3[3,2] / A3[2,2]
  matrix_E32 <- matrix(c(1,0,0,0,1,0,0,E32,1), nrow=3, byrow = T)
  U <- matrix_E32 %*% A3
 
  L <- solve(matrix_E21) %*% solve(matrix_E31) %*% solve(matrix_E32)

  A <- L %*% U

  print(A) # Matrix A.
  print(L) # Lower Trangular Matrix.
  print(U) #Upper Trangular Matrix.
 
  #Testing
  (L %*% U == A)}

```