---
title: 'DATA 605: Assignment 3'
author: "MUSA T. GANIYU"
date: "September 17, 2016"
output:
  pdf_document: default
  html_document: default
---
> QUESTION 1.1:

+ What is the rank of the matrix A?

```{R}
A1 <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow = 4, byrow = T)

A1
dim(A1)

```

Solution:
From the above matrix, its known that its dimension is 4x4(a square matrix), therefore it rank is 4



> QUESTION 1.2:

+ Given an mxn matrix where m > n, what can be the maximum rank? The mini-
mum rank, assuming that the matrix is non-zero?


Solution:

Since the rank is the number of all non-zero row, the
rank has to be no greater than the smaller of the row or column dimension is n.


> QUESTION 1.3 

+ What is the rank of Matrix B?

```{R}

B <- matrix(c(1,2,1,3,6,3,2,4,2), nrow = 3, byrow = T)

B

dim(B)



R1 <- B[1, ]
R2 <- B[2, ]
R3 <- B[3, ]

a <- R1-(1/3)%*%R2
b <- R3-(2/3)%*%R2


Mat <- matrix(c(a,b,R2), nrow = 3, byrow = T)

Mat

```

Solution:

Since the rank is the number of all non-zero row, therefore the rank is 1.



> QUESTION 2:Compute the eigenvalues and eigenvectors of the matrix A.

```{r}


A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow = 3, byrow = T)

A


I <- matrix(c("x" ,0,0,0,"x" ,0,0,0,"x" ), nrow = 3, byrow = T)

I
```


$det\left( \left[ \begin{matrix} 1 & 2 & 3 \\ 0 & 4 & 5 \\ 0 & 0 & 6 \end{matrix} \right] \quad -\quad \left[ \begin{matrix} \lambda  & 0 & 0 \\ 0 & \lambda  & 0 \\ 0 & 0 & \lambda  \end{matrix} \right]  \right) \quad =\quad det\left( \left[ \begin{matrix} 1-\lambda  & 2 & 3 \\ 0 & 4-\lambda  & 0 \\ 0 & 0 & 6-\lambda  \end{matrix} \right]  \right)$

$=\quad (1-\lambda )\left[ (4-\lambda )(6-\lambda )\quad -(0*0) \right] \quad -\quad (2)\left[ (0)(6-\lambda )-(0*0 \right] \quad +\quad (3)\left[ (0*0)-(0*4-\lambda ) \right]$ 



$=\quad (1-\lambda )\left[ (4-\lambda )(6-\lambda ) \right] \quad =\quad 0$



+ From the above, we can deduce (solve algebraically) that $\lambda$ are : 1, 4 and 6



Their respective Eigenvalues are:

For $\lambda_1$ = 1, its eigenvectors are $\left[\begin{matrix} 1.0000 \\ 0.0000 \\ 0.0000 \end{matrix} \right]$

For $\lambda_2$ = 4, its eigenvestors are $\left[\begin{matrix} 1.6000 \\ 2.5000 \\ 1.0000 \end{matrix} \right]$

For $\lambda_3$ = 6, its eigenvectors are $\left[\begin{matrix} 0.6667 \\ 1.0000 \\ 0.0000 \end{matrix}  \right]$



**OR**

```{r}

eigen(A, only.values = FALSE, EISPACK = TRUE)

```